### 2018-01-14 Paper/PaperReport/PaperReport.tex

\documentclass[12pt]{report}
\usepackage[a4paper]{geometry}
\usepackage[myheadings]{fullpage}
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage{graphicx, wrapfig, subcaption, setspace, booktabs}
\usepackage[T1]{fontenc}
\usepackage[font=small, labelfont=bf]{caption}
\usepackage{fourier}
\usepackage[protrusion=true, expansion=true]{microtype}
\usepackage[english]{babel}
\usepackage{sectsty}
\usepackage{url, lipsum}
\usepackage{listings}
\usepackage{float}
\usepackage{todonotes}
\graphicspath{{images/}}

\lstset{
	basicstyle=\ttfamily,
	columns=fullflexible,
	frame=single,
	breaklines=true,
	postbreak=\mbox{\textcolor{red}{$\hookrightarrow$}\space},
}

\newcommand{\HRule}[1]{\rule{\linewidth}{#1}}
\onehalfspacing
\setcounter{tocdepth}{5}
\setcounter{secnumdepth}{5}

%-------------------------------------------------------------------------------
% HEADER & FOOTER
%-------------------------------------------------------------------------------
\pagestyle{fancy}
\fancyhf{}
\setlength\headheight{15pt}
\fancyfoot[R]{Page \thepage\ of \pageref{LastPage}}
%-------------------------------------------------------------------------------
% TITLE PAGE
%-------------------------------------------------------------------------------

\begin{document}

	\title{ \normalsize \textsc{Large-Scale Data Analysis Techniques}
		\\ [2.0cm]
		\HRule{0.5pt} \\
		\LARGE \textbf{\uppercase{A Review on Multi-Label Learning Algorithms}}
		\HRule{2pt} \\ [0.5cm]
		\normalsize \today \vspace*{5\baselineskip}}

	\date{}
	\author{
		Nikiforos Pittaras, M1422\\
		Chrysoula Themeli, M1423 \\ 
		National and Kapodistrian University of Athens\\
		Department of Informatics and Telecommunications }

	\maketitle
	\tableofcontents
	\listoffigures
	\newpage

	%-------------------------------------------------------------------------------
	% Section title formatting
	\sectionfont{\scshape}
	%-------------------------------------------------------------------------------

	%-------------------------------------------------------------------------------
	% Summary of the paper
	%-------------------------------------------------------------------------------

	\section*{Summary}
	\addcontentsline{toc}{section}{Summary}

	%\begin{figure}[H]
	%	\centering
	%	\includegraphics[width=0.8\textwidth]{semsearch.png}
	%	\caption{An overview of the SemSearch architecture}
	%	\centering
	%\end{figure}

	%-------------------------------------------------------------------------------
	% REFERENCES
	%-------------------------------------------------------------------------------
	\newpage
	\addcontentsline{toc}{section}{Bibliography}
	\begin{thebibliography}{10}
		\bibitem{intelligent} 
		Madhu, G., Dr A. Govardhan, and Dr TV Rajinikanth. "Intelligent semantic web search engines: a brief survey." \url{https://arxiv.org/pdf/1102.0831.pdf}
	\end{thebibliography}
\end{document}