### 2018-01-20 Paper/PaperReport/PaperReport.tex
Update project report

\documentclass[12pt]{report}
\usepackage[a4paper]{geometry}
\usepackage[myheadings]{fullpage}
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage{graphicx, wrapfig, subcaption, setspace, booktabs}
\usepackage[T1]{fontenc}
\usepackage[font=small, labelfont=bf]{caption}
\usepackage{fourier}
\usepackage[protrusion=true, expansion=true]{microtype}
\usepackage[english]{babel}
\usepackage{sectsty}
\usepackage{url, lipsum}
\usepackage{listings}
\usepackage{float}
\usepackage{todonotes}
\usepackage{stmaryrd}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{bm}
\graphicspath{{images/}}

\lstset{
	basicstyle=\ttfamily,
	columns=fullflexible,
	frame=single,
	breaklines=true,
	postbreak=\mbox{\textcolor{red}{$\hookrightarrow$}\space},
}

\usepackage{hyperref}
\hypersetup{
	colorlinks,
	citecolor=black,
	filecolor=black,
	linkcolor=black,
	urlcolor=black
}

\newcommand{\HRule}[1]{\rule{\linewidth}{#1}}
\onehalfspacing
\setcounter{tocdepth}{5}
\setcounter{secnumdepth}{5}

%-------------------------------------------------------------------------------
% HEADER & FOOTER
%-------------------------------------------------------------------------------
\pagestyle{fancy}
\fancyhf{}
\setlength\headheight{15pt}
\fancyfoot[R]{Page \thepage\ of \pageref{LastPage}}
%-------------------------------------------------------------------------------
% TITLE PAGE
%-------------------------------------------------------------------------------

\begin{document}

	\title{ \normalsize \textsc{Large-Scale Data Analysis Techniques}
		\\ [2.0cm]
		\HRule{0.5pt} \\
		\LARGE \textbf{\uppercase{A Review on Multi-Label Learning Algorithms}}
		\HRule{2pt} \\ [0.5cm]
		\normalsize \today \vspace*{5\baselineskip}}

	\date{}
	\author{
		Nikiforos Pittaras, M1422\\
		Chrysoula Themeli, M1423 \\ 
		National and Kapodistrian University of Athens\\
		Department of Informatics and Telecommunications }

	\maketitle
	\tableofcontents
	\listoffigures
	\newpage

	%-------------------------------------------------------------------------------
	% Section title formatting
	\sectionfont{\scshape}
	%-------------------------------------------------------------------------------

	%-------------------------------------------------------------------------------
	% Summary of the paper
	%-------------------------------------------------------------------------------

	\section*{Introduction}
	\addcontentsline{toc}{section}{Introduction}

	\subsection*{Multi-label learning}
	\addcontentsline{toc}{subsection}{Multi-label learning}
	The paper "A Review on Multi-Label Learning Algorithms" analyzes the
  multi-label learning field in supervised classification. Supervised
  classification trains a model $f: X \rightarrow Y$ and for each pair
  $f(x_i,y_i)$, it outputs a real number $r$ that represents the confidence that
  label $y_i$ characterizes $x_i$. Given $r$ and a thresholding scheme, labels
  are marked as relevant or not relevant to the input data vector $x_i$ (e.g.
  for $r$ representing probability estimates, $r \in [0,1]$, a simple
  thresholding function is $t(x_i) = 0.5$. In other words, labels with outputs
  greater than $0.5$ are assigned to $x_i$ by our model). Standard, single-label learning maps an
  example to a single label, completely ignoring multiple semantics that
  real-objects usually have. On the other hand, multi-label learning allows an
  example/instance $x_i$ to be mapped to more than one label, i.e. to any label
  vector which is a subset of $Y$. However, this results to an exponentially growing search space, since
  the latter depends on the size of the powerset of $Y$, 
  $|\mathbb{P(Y)}| = 2^{|Y|}$.

	One straightforward solution is to exploit label correlations. Real objects
  are usually characterized by similar concepts. For example, if an article is
  characterized by label "football" it follows that it will be also
  characterized by the term "sports" and it is likely that it will be labeled
  with ``crowd'' or ``entertainment''. On the other hand, the label "science"
  will be likely to be irrelevant. In order to exploit such kind of label
  correlations, three different algorithm strategies are examined:

	\begin{itemize}
		\item \textbf{First-order strategies: }These algorithms ignore label
      correlations and perform a "one-vs-all" strategy. They often transform a
      multi-label problem into multiple single-label problems, combining the
      result in an aggregation-like manner. Algorithms of this category are
      simple, scalable, have the capability of parallel implementation, but
      offer suboptimal performance.
		\item \textbf{Second-order strategies: }Algorithms in this category consider
      \emph{pairwise} label correlations, i.e. tuples of labels. Having a good
      trade-off between generalization performance and scalability, they
      generally outperform first-order strategies but their limited exploration
      of the label dependency space leads to lacking performance in some real
      world applications.
		\item \textbf{High-order strategies: }These algorithms capture a high degree
      of label correlations, having a large comlexity and strong modeling
      capabilities. This of course leads to increased performance but renders them computationally demanding and less scalable.
	\end{itemize}

	\subsection*{Evaluation metrics}
	\addcontentsline{toc}{subsection}{Evaluation metrics}

	With respect to performance evaluation, traditional single-label evaluation metrics are extended to
  incorporate the multi-labeled output. This extension is categorizable into two
  groups. One with respect to the expansion ``axis'' (across examples or labels)
  and the task type (classification or ranking).

    Before proceeding with the metrics, true positive(TP), false positive(FP),
    true negative (TN) and false negative (FN) of label $y_j$ will be defined:
		\begin{itemize}
			\item \emph{True positive: } $TP_j = |\{x_i | y_j \in Y_i \wedge y_j \in h(x_i), 1 \leq i \leq p \}|$
			\item \emph{False positive: }$FP_j = |\{x_i | y_j \notin Y_i \wedge y_j \in h(x_i), 1 \leq i \leq p \}|$
			\item \emph{True negative: }$TN_J = |\{x_i | y_j \notin Y_i \wedge y_j \notin h(x_i), 1 \leq i \leq p \}|$
			\item \emph{False negative: }$FN_J = |\{x_i | y_j \in Y_i \wedge y_j \notin h(x_i), 1 \leq i \leq p \}|$
		\end{itemize}

	\begin{itemize}
		\item \textbf{Example-based: }This kind of metrics evaluate multi-labeled performance on each example and then the result is spread to the whole dataset. The metrics are divided into two perspectives:
      \todo{Form them up into a table}
		\begin{itemize}
			\item Classification:
        \begin{table}
          \centering
          \begin{tabular}{cc} 
          \toprule
            one & two \\
          \midrule
            one & two \\
            one & two \\
          \bottomrule
          \end{tabular}
          \caption{Evaluation metrics for multi-label learning}
          \label{label:evaluation_metrics}
        \end{table}

			\begin{itemize}
				\item Precision: $P(h) = \frac{1}{p} \sum_{i=1}^{p} \frac{|Y_i \cap h(x_i)|}{h(x_i)|}$
				\item Recall: $R(h) = \frac{1}{p} \sum_{i=1}^{p} \frac{|Y_i \cap h(x_i)|}{|Y_i}$
        \item Subset Accuracy:  $SA(h) = \frac{1}{p} \sum_{i=1}^p \llbracket h(x_i) = Y_i \rrbracket$
				\item Hamming Loss: $H(h) = \frac{1}{p} \sum_{i=1}^{p} [h(x_i)\Delta Y_i] $
				\item Accuracy: $A(h) = \frac{1}{p} \sum_{i=1}^{p} \frac{|Y_i \cap h(x_i)|}{|Y_i \cup h(x_i)|}$
				\item $F^ \beta$ -metric:  $F^ \beta _{exam} (h) = \frac{(1+ \beta ^2) \cdot Precision_{exam}(h) \cdot Recall_{exam}(h)}{\beta ^2 \cdot Precision_{exam}(h) + Recall_{exam}(h)}$
			\end{itemize}
			\item Ranking:
			\begin{itemize}
				\item One-error: $one-error(f) = \frac{1}{p} \llbracket [argmax_{y \in Y}f(x_i,y)] \notin Y_i \rrbracket$
				\item Coverage:  $coverage(f) = \frac{1}{p} \sum_{i=1}^{p} max_{y \in Y_i} rank_f(x_i,y) - 1$
				\item Ranking loss: $rloss(f) = \frac{1}{p} \sum_{i=1}^{p} \frac{1}{|Y_i| {\overset{\sim}{|Y_i|}}} |\{(y', y'') | f(x_i,y') \leq f(x_i,y''), (y', y'') \in Y_i \times {\overset{\sim}{|Y_i|}} \}| $ 
				\item Average Precision:  $agvprec(f) = \frac{1}{p} \sum_{i=1}^{p} \frac{1}{|Y_i|} \sum_{y \in Y_i} \frac{|\{y' | rank_f(x_i,y') \leq rank_f(x_i,y), y' \in Y_i\}|}{rank_f(x_i,y)}$
			\end{itemize}
		\end{itemize}
  \item \textbf{Label-based: }
		The label-based metrics are also separated in two perspectives:
		\begin{itemize}
			\item Classification: Let $B(TP_j, FP_j, TN_J, FN_J)$ represent some specific binary classification metric $(B \in \{Accuracy, Precision, Recall, F^ \beta \})$. The metrics from classification perspective are the following:
			\begin{itemize}
				\item Macro-averaging: $B_{macro}(h) = \frac{1}{q} \sum_{1}^{q} B(TP_j, FP_j, TN_J, FN_J)$
				\item Micro-averaging: $B_{micro}(h) = B(\sum_{1}^{q}TP_j, \sum_{1}^{q}FP_j, \sum_{1}^{q}TN_j, \sum_{1}^{q}FN_j)$
			\end{itemize}
			\item Ranking:
			\begin{itemize}
				\item \emph{AUC macro: }$AUC_{macro} = \frac{1}{q} \sum_{1}^{q} \frac{|\{(x', x'') | f(x',Y_j) \geq f(x'',Y_j), (x',x'') \in Z_j \times \overset{\sim}{|Z_j|} \}|}{|Z_j|{\overset{\sim}{|Z_j|}}}$
				\item \emph{AUC micro: }$AUC_{micro} = \frac{|\{(x',x'',y',y'') | f(x',y') \geq f(x'',y''), (x',y') \in S^+, (x'',y'') \in S^- \}|}{|S^+||S^-|}$
			\end{itemize}
		\end{itemize} 
	\end{itemize}

	\section*{Multi-label learning algorithms}
	\addcontentsline{toc}{section}{Multi-label learning algorithms}
	In this paper, eight representative multi-label algorithms are presented and analyzed, seleceted with respect to below criteria:
	\begin{itemize}
		\item[$\checkmark$] Broad, noteworthy and/or unique characteristics
		\item[$\checkmark$] Primitive impact (i.e. leads to a number follow-up related methods)
		\item[$\checkmark$] Favorable influence (popular and/or highly-cited) 
	\end{itemize}

	The algorithms to be presented are grouped in two categories:
	\begin{itemize}
		\item \textbf{Problem Transformation Methods: }Algorithms of this category
      transform the multi-labeled problem into into other well-established
      single-labeled learning scenarios, fusing the results into a multi-label output (``Fit data to algorithm'' philosophy)
		\item \textbf{Algorithm Adaptation Methods: }Here, algorithms adapt popular learning techniques to deal with multi-label data directly ("Fit algorithm to data" philosophy)
	\end{itemize}

	\subsection*{Problem Transformation Methods}
  We move on to present the first category of methods, which transform the
  multi-label problem into a set of single label methods. See figure \ref{figure:ptm} for an overview of the approaches examined.

	\addcontentsline{toc}{subsection}{Problem Transformation Methods}
	\begin{figure}[H]
		\centering
		\includegraphics[width=0.9\textwidth]{pt.png}
		\caption{Problem Transformation Algorithms}
		\centering
    \label{figure:ptm}
	\end{figure}

	\subsubsection*{Binary Relevance}
	\addcontentsline{toc}{subsubsection}{Binary Relevance}
	Binary relevance (BR) decomposes the multi-learning problem into $q$
  independent binary classification problems, each of them corresponding to a
  possible label in the label-set $Y$. The first task is to construct a
  corresponding binary training set for each label, on which a binary learning
  algorithm is used in order to train the binary classifier that discriminates
  class $y_j$, i.e. $h_j(\cdot)$. Each training instance is involved in the
  learning process of all $q$ classifiers(cross-training strategy), which is
  regarded as positive instance in case of relevant labels and negative
  otherwise. For new instances, the algorithm predicts the relevant labelset by
  querying each binary classifier and combining the relevant labels. A label is
  considered relevant according to the thresholding scheme adopted in the classifier.

	This is a first-order approach algorithm, since it ignores completely any
  label associations. However, since an one-vs-rest scheme is used and
  classifiers are built for each label separately, the algorithm has the
  capability of parallel implementation. Finally, this separation renders the
  algorithm sensitive to class-imbalance in the dataset\footnote{The case where
    the number of positives and negatives for a class differs significantly}.

	\subsubsection*{Classifier Chains}
	\addcontentsline{toc}{subsubsection}{Classifier Chains}
	The logic of this algorithm is to transform the multi-label learning problem
  into a chain of binary classification problems, using subsequent binary
  classifiers in a chain, built upon the predictions of preceding ones and
  ordered by a permutation function $f_p$. First of all, a binary training set
  is constructed by enriching each instance with the confidence of preceding
  classifiers, which are concatenated with the data vector $x_i$. Moreover, a binary algorithm is used in step $i$ to
  determine whether label $y_i$ label is relevant or not for an instance. For
  unknown instances the algorithm predicts the relevant labels by iteratively
  traversing the classifier chain producing thresholded confidence scores ($\lambda _{\tau(j)}^x \in {-1, +1}$ in the
  same order as within the training process by $f_p$ represents the predicted binary assignment of $y_{\tau(j)}$)

	This is a high-order method, that exploits label correlations to a degree, but in a random manner. However, it is highly sensitive to the permutation function, since the classifiers' ordering directly affects the result. One other disadvantage is that iterative operation prevents parallel implementation. Finally, one possible improvement to overcome the affect of classifiers' ordering is to use ensemble schemes.

	\subsubsection*{Claibrated Label Ranking}
	\addcontentsline{toc}{subsubsection}{Claibrated Label Ranking}
	This algorithm transforms the learning problem into a label pairwise
  comparison ranking problem. For $q$ labels, $q(q-1)/2$ binary classifiers are
  constructed to discriminate between each label pair via a binary
  classification algorithm $h_{jk}(x)$. Firstly, respective pairwise training
  sets $D_{jk}$ are constructed, where $D_{jk}: \{x_i, Y_i | y_j \in Y_i \oplus
  y_k \in Y_i\}$, i.e. each $D_{jk}$ contains examples either in class $j$ or
  $k$. The learning system votes for each example and if $h_{jk}(x)>0$, then
  $x_i$ is associated with label $y_j$, otherwise $y_j$ is irrelevant and $x_i$
  is assigned to $y_k$. For unknown instances, all classifiers' votes are
  aggregated and ranked. In addition, a virtual label $y_v$ is used as an
  artificial splitting point between relevant and irrelevant labels, serving as
  a decision threshold for label relevance.

	This is a second-order approach algorithm using a one-vs-one scheme. The
  advantage of this method is that it smooths out the class-imbalance problem
  (where the number of a class positives is much less than the other's class
  positives) by operating on label pairs. The disadvantage is that the number of
  classifiers is quadratic to $|Y|$, compared to linear for Binary Relevance or
  Classifier Chains algorithms. However, the performance can be improved by
  using pruning methods to reduce the search space.

	\subsubsection*{Random k-Labelsets}
	\addcontentsline{toc}{subsubsection}{Random k-Labelsets}
	The random k-labelsets algorithm transforms the multi-label learning problem
  into an ensemble of multi-class classification problems. Each such partition
  targets a random subset of $\mathbb{P}(Y)$, limited to subsets present in the
  training dataset $X$, classifying it using Label Powerset (LP) techniques. In
  more detail, the multi-label dataset is transformed into single-label by
  treating each distinct labelset as a new class. Thus, each training example
  mapped to a new single class and classified through some single-label
  classification method. An instance $x_i$ is assigned to label $y_i$ when the votes received for $y_i$ from the ensemble exceed half the max possible that this label can get. For unseen instances $x$, LP predicts its associated label set $Y$ by querying the prediction of multi-class classifier and then mapping it back to the power set of $Y$.

	This is a high-order approach algorithm since it utilizes a variety of label correlations. However, there are two limitations in terms of practical feasibility: 
	\begin{itemize}
		\item[$\diamond$] \emph{Incompleteness: }The algorithm is data-sensitive and cannot generalize to labelsets that do not appear into the training set.
		\item[$\diamond$] \emph{Inefficiency: }A large label set $Y$ implies high training complexity and extremely few training examples for some newly mapped classes.
	\end{itemize}

	A simple but powerful modification is the Random k-Labelsets variant, which
  runs an ensemble of executions, each with $k$-sized label sets. The ensemble
  scheme imporives on the predictive completeness of the system, while the
  limited labelset size guarantees computational efficiency. Additionally,
  pruning distinct labelsets which are less frequent than a lower threshold is
  another strategy to improve the method's performance.

	\subsection*{Algorithm Adaptation Methods}
  The next category refers to algorithm adaptation methods, i.e. strategies that
  directly address the multi-label nature of the problem. See figure
  \ref{figure:aam} for an overview of the approaches examined.

	\addcontentsline{toc}{subsection}{Algorithm Adaptation Methods}

	\begin{figure}[H]
		\centering
		\includegraphics[width=0.9\textwidth]{aa.png}
		\caption{Algorithm Adaptation Methods}
		\centering
    \label{figure:aam}
	\end{figure}

	\subsubsection*{Multi-Label k-Nearest Neighbor (ML-kNN)}
	\addcontentsline{toc}{subsubsection}{Multi-Label k-Nearest Neighbor (ML-kNN)}
	The ML-kNN algorithm adapts nearest neighbor techniques to the multi-label
  setting by utilizing the maximum a posteriori (MAP) rule, to make a prediction
  by reasoning with the labeling information embodied in the neighbors. For new
  instances $x$, $N(x)$ represents the set of its $k$ nearest neighbors
  identified in the dataset by a specified similarity or distance measure. Let
  $H_j \equiv (y_j \in Y_i)$, $C_j = \sum_{x_k \in N(x_i)} [\delta(y_j \in
  Y_k)]$, where $H_j$ is the event that $\textbf{x}$ has the label $y_j$ and
  $C_j$ records the number of x’s neighbors with label $y_j$. Moreover,
  $P(H_j|C_j)$ is the posterior probability that $H_j$ holds in the condition
  that $\textbf{x}$ has exactly $C_j$ neighbors. The posteriors $P(H_j|C_j)$ are
  calculated using Bayes' theorem with respect to the priors and the likelihoods.

	The prior probabilities ($P(H_j)$, $P(\neg H_j)$) are computed via a frequency counting strategy, where a smoothing parameter $s$ is used to control the effect of uniform prior on the estimation (usually $s=1$). In addition, likelihoods $P(C_j|H_j)$ and $P(C_j| \neg H_j)$ are computed using two frequency arrays:
	\begin{itemize}
		\item $\kappa(r)$, the number of examples labelled $y_j$ with $r$ neighbours labelled $y_j$
		\item $\overset{\sim}{\kappa}(r)$, the number of examples not labelled $y_j$ with $r$ neighbours labelled $y_j$
	\end{itemize}

	This algorithm is a first-order approach that uses Bayesian reasoning. Being a
  nearest-neighbour method, the decision boundary can be modified on-line when
  new instances appear. Moreover, the use of prior probabilities can amend the
  class imbalance issue to a degree. A number of extensions have been proposed that take into account label associations.

	\subsubsection*{Multi-Label Decision Tree (ML-DT)}
	\addcontentsline{toc}{subsubsection}{Multi-Label Decision Tree (ML-DT)}
	This method adapts decision tree techniques for multi-label classification.
  Using the multi-label entropy measure, decision tree is built recursively by
  splitting features that yield the maximum information gain. Let $T = \{(x_i,
  Y_i) | 1 \leq i \leq n \}$ be the multi-label dataset with n examples. The
  algorithm splits the $x$ vector at the feature $l$ that maximizes the
  information gain criterion, where that value in $x$ is  $\theta$. This
  partitions the dataset $T$ into branches ($T^-$ \& $T^+$), where $x_{il} \leq
  \theta, x_i \in T^-$ and $x_{il} > \theta, x_i \in T^+$. This process is
  invoked recursively by treating each time one of the two branches as the new
  root node and terminates until some stopping criterion $C$ is met (e.g.
  subtree size). Each label subset is treated as a new class and single-label entropy is computed. Any new instances are assigned the label of the majority of members of the leaf that they arrive.

	In order to ensure low computational cost and high efficiency, the algorithm
  assumes independence among the labels; thus, it is a first-order approach.
  Extensions are proposed to reduce computational cost and exploit label interrelations, such as pruning and ensemble strategies.

	\subsubsection*{Ranking Support Vector Machine (Rank-SVM)}
	\addcontentsline{toc}{subsubsection}{Ranking Support Vector Machine (Rank-SVM)}
	Rank-SVM adapts the maximum margin strategy to the multi-label classification
  problem. $q$ linear classifiers are optimized to minimize the empirical
  ranking loss. The learning system is composed by classifiers that distinguish
  between relevant and irrelevant pairs. Each classifier is parameterized with $W = \{w_j,b_j | 1 \leq j \leq q\}$, where $w_j \in \mathbb{R}^d$ and $b_j \in \mathbb{R}$ are the weight vector and the bias
	for the j-th class label $y_j$ respectively. Rank-SVM defines the system’s
  margin on $(x_i,Y_i)$ by considering its ranking ability on the relevant and
  irrelevant label pairs of $x_i$ $(y_j,y_k) \in Y_i \times \bar Y_i$, which are
  separated by a function of their classifiers $h_j, h_k$: $\left<w_j-w_k, x_i
  \right> + b_j - b_k=0$. The margin is defined as the minimum distance from all
  related-unrelated pairs and is minimized along with the training loss
  function. At consistent training, the classifier will return positive margin.
  For new test instances, the algorithms retrieves a ranked list of label pairs
  according to the signed distance to the decision hyperplane.

	Non-linear problems can be managed by exploiting kernel feature mapping via
  the kernel trick, which is popular in SVM methods. The optimization method is
  a convex quadratic programming problems and can thus be solved by any QP
  solver.
  This algorithm is a second-order approach using pairwise label associations.
  It is adaptable to a variety of tasks and data by using alternative loss
  functions and kernel feature mapping. Kernel selection can be accomplished by
  multiple kernel learning (MKL) techniques.

  \subsubsection*{Collective Multi-Label Classifier}
	 \addcontentsline{toc}{subsubsection}{Collective Multi-Label Classifier}
   The CML method is a multi-label variant of Conditional Random Field methods.

	 This algorithm involves the maximum entropy principle, where correlations
   among labels are encoded as binary random vectors into a joint probability
   distribution function and label correlations are encoded into the
   distribution as constraints.
   The conditional $p(\bm{y} | x)$ is computed via the maximum entropy
   criterion, where the information entropy measure $H_p(x,\bm{y})$ is maximized subject to each constraint.
   The latter are modeled as $ \mathbb{E} [ f_k(x,\bm{y}) ] = F_k  $, where $F_k$ is
 learned from the traininng data.
 The optimal solution is arrived at via Lagrange multipliers, assuming gaussian
 priors for the variables involved, and has the form of a Gibbs probability
 distribution.
 Test instances are labelled with respect to $\text{argmax}_y\text{ }p(\bm{y} | x)$.

   This is a second-order approach, where all label pairs are considered and not
   just the relevant-irrelevant ones which was the Rank-SVM approach. It is a
   convex constraint optimization problem, and is thus solvable by off the shelf
   CP solvers. The argmax operation during inference is expensive, which renders
   the method very computationally inefficient for large label spaces, if
   pruning is not applied. Varying the expression form of the constraints gives rise to difference
   version of the CML method.

	 \subsection*{Summary of Multi-Label Learning Algorithms}
	 \addcontentsline{toc}{subsection}{Summary of Multi-Label Learning Algorithms}

	 \begin{figure}[H]
	 	\centering
	 	\includegraphics[width=1\textwidth]{summary.png}
	 	\caption{Summary of Multi-Label Learning Algorithms}
	 	\centering
	 \end{figure}

 	\section*{Related Learning Methods}
 	\addcontentsline{toc}{section}{Related Learning Methods}
 	Four related learning methods are mentioned in this paper, which can be found below:
 	\begin{enumerate}
 		\item \textbf{Multi-instance learning: }This method studies the problem
      where each example is described by a bag of instances while associated
      with a single binary label. The bag is assigned a positive label in the
      case of at least one positively labelled member. It models complex semantics of $x_i$ in input space rather than its output.
 		\item \textbf{Ordinal classification: } This scheme assumes label relevance
      to an instance is not binary but graded, producing a vector of ordinal graded
      membership of $x$ to a label. Multi-label classification is induced via a
      transformation of the original problem into a set of ordinal problems.
 		\item \textbf{Multi-task learning: }This method studies the problem where
      multiple tasks are trained in parallel. Tasks can share information, and
      knowledge from related tasks is used as an inductive bias to help improve
      the generalization performance of other tasks. The  tasks can share a
      common feature space or use different representations, as is common in
      multi-label learning. In this setting, it is reasonable to have small task
      workloads, since the execution is of a parallel nature. 
 		\item \textbf{Data streams classification: } Here, data is generated in real
      time, corresponding to the real-time nature of some task. Apart from the
      tight temporal constraints, the defining challenge here is the concept
      drift problem, where the concept classified can change over time. A common
      strategy to this effect is to gradually discount the significance of past
      data in favor of newly arrived batches.

 	\end{enumerate}

	\section*{Conclusion}
	\addcontentsline{toc}{section}{Conclusion}
	Summarizing, this paper defines the multi-label learning problem, presenting 8
  multi-label learning representative algorithms, mentions some additional
  related learning settings. Some important future goals are the formal
  characterization on the underlying concept and / or mechanism on the
  appropriate usage of label correlations (especially on large output spaces).
  In addition, thorough experimental comparative studies are essential to
  discover the advantages and disadvantages of different multi-label learning
  algorithms, so as to aid the conquest of the aforementioned theoretical goal.
  Some online resources on multi-label learning are presented on figure \ref{figure:resources}.

	\begin{figure}[H]
		\centering
		\includegraphics[width=1\textwidth]{online.png}
		\caption{Online resources for Multi-Label Learning}
    \label{figure:resources}
	\end{figure}

	%-------------------------------------------------------------------------------
	% REFERENCES
	%-------------------------------------------------------------------------------
	\addcontentsline{toc}{section}{Related Work and Bibliography}
	\renewcommand\bibname{Related Work and Bibliography}
	\begin{thebibliography}{9}
		\bibitem{comp} 
		Madjarov, Gjorgji, et al. "An extensive experimental comparison of methods for multi-label learning." Pattern Recognition 45.9 (2012): 3084-3104
		\bibitem{overview} Tsoumakas, Grigorios, and Ioannis Katakis. "Multi-label classification: An overview." International Journal of Data Warehousing and Mining 3.3 (2006)
		\bibitem{correlations} Zhang, Min-Ling, and Kun Zhang. "Multi-label learning by exploiting label dependency." Proceedings of the 16th ACM SIGKDD international conference on Knowledge discovery and data mining. ACM, 2010

\bibitem{trees} Vens, Celine, et al. "Decision trees for hierarchical multi-label classification." Machine Learning 73.2 (2008): 185-214.
\bibitem{rkl} Tsoumakas, Grigorios, and Ioannis Vlahavas. "Random k-labelsets: An ensemble method for multilabel classification." Machine learning: ECML 2007 (2007): 406-417.
\bibitem{mlknn} Zhang, Min-Ling, and Zhi-Hua Zhou. "ML-KNN: A lazy learning approach to multi-label learning." Pattern recognition 40.7 (2007): 2038-2048.
\bibitem{rksvm} Elisseeff, André, and Jason Weston. "A kernel method for multi-labelled classification." Advances in neural information processing systems. 2002.
\bibitem{cml}  Ghamrawi, Nadia, and Andrew McCallum. "Collective multi-label classification." Proceedings of the 14th ACM international conference on Information and knowledge management. ACM, 2005.
\bibitem{ordinals}  Cheng, Weiwei, Eyke Hüllermeier, and Krzysztof J. Dembczynski. "Graded multilabel classification: The ordinal case." Proceedings of the 27th international conference on machine learning (ICML-10). 2010.

	\end{thebibliography}
\end{document}